### 2024-01-05 tex/figures/timeline_mfrr_tikz.tex
Preparing for submission to new journal.

\documentclass[crop,tikz]{standalone}
\usepackage[usenames]{color} %used for font color
\usepackage{amssymb} %maths
\usepackage{amsmath} %maths
\usepackage[utf8]{inputenc} %useful to type directly diacritic characters
\usepackage{tikz}
\usetikzlibrary{calc,arrows,patterns,intersections}
\usetikzlibrary{shapes.multipart}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}
\begin{document}
\definecolor{RectangleColor}{RGB}{197,224,180}

\begin{tikzpicture}[scale=0.75,every text node part/.style={align=center}]

  \draw[rounded corners, fill=RectangleColor, very thick] (0, 1) rectangle (4, 3) node[midway]{\LARGE mFRR \\ \LARGE reservation};
  \node at (2,0) {\LARGE $\lambda_{h}^{\rm{r}} p_{h}^{\rm{r},\uparrow}$};
  \node at (2,-2) {\LARGE D$-$1};
  \node at (2,-2.7) {\large (9:30am)};

  \draw[rounded corners, fill=RectangleColor, very thick] (5, 1) rectangle (9, 3) node[midway]{\LARGE Day-ahead \\ \LARGE cost};
  \node at (7,0) {\LARGE $-\lambda_{h}^{\rm{s}} p_{h}^{\text{Base}}$};
  \node at (7,-2) {\LARGE D$-$1};
  \node at (7,-2.7) {\large (12:00pm)};

  \draw[rounded corners, fill=RectangleColor, very thick] (10, 1) rectangle (14, 3) node[midway]{\LARGE Regulating \\ \LARGE power bid};
  \node at (12,0) {\LARGE $\lambda_{h}^{\text{bid}}$};
  \node at (12,-2) {\LARGE D$-$1};
  \node at (12,-2.7) {\large (5:00pm)};

  \draw[name path=VerticalDashedLine, dashed, line width=0.45mm, black](14.5,-1)--(14.5,3);

  \draw[rounded corners, fill=RectangleColor, very thick] (15, 1) rectangle (19, 3) node[midway]{\LARGE mFRR \\ \LARGE activation};
  \node at (17,0) {\LARGE $\lambda_{h}^{\rm{b}} p_{h}^{\rm{b},\uparrow}$};
  \node at (17,-2) {\LARGE D};
  \node at (17,-2.7) {\large (12pm -- 12am)};

  \draw[rounded corners, fill=RectangleColor, very thick] (20, 1) rectangle (24, 3) node[midway]{\LARGE Rebound \\ \LARGE cost};
  \node at (22,0) {\LARGE $-\lambda_{h}^{\rm{b}} p_{h}^{\rm{b},\downarrow}$};
  \node at (22,-2) {\LARGE D};
  \node at (22,-2.7) {\large (12pm -- 12am)};

  % Right arrow
  \node (source) at (0, -1) {};
  \node (destination) at (24,-1){};
  \draw[->, line width=0.65mm, black](source)--(destination);

\end{tikzpicture}

\end{document}