### 2024-01-05 tex/figures/timeline_mfrr_variables_tikz.tex

\documentclass[crop,tikz]{standalone}
\usepackage[usenames]{color} %used for font color
\usepackage{amssymb} %maths
\usepackage{amsmath} %maths
\usepackage[utf8]{inputenc} %useful to type directly diacritic characters
\usepackage{tikz}
\usetikzlibrary{calc,arrows,patterns,intersections}
\usetikzlibrary{shapes.multipart}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}
\begin{document}
\definecolor{RectangleColor}{RGB}{197,224,180}

\begin{tikzpicture}[scale=0.75,every text node part/.style={align=center}]

  \draw[rounded corners, fill=RectangleColor, very thick] (0, 1) rectangle (4, 3) node[midway]{\LARGE mFRR \\ \LARGE reservation};
  \node at (2,0) {\LARGE  $p_{h}^{\rm{r},\uparrow}$};
  \node at (2,-2) {\LARGE D$-$1};
  \node at (2,-2.7) {\large (9:30am)};

  \draw[rounded corners, fill=RectangleColor, very thick] (5, 1) rectangle (9, 3) node[midway]{\LARGE Day-ahead \\ \LARGE clearing};
  \node at (7,-2) {\LARGE D$-$1};
  \node at (7,-2.7) {\large (12:00pm)};

  \draw[rounded corners, fill=RectangleColor, very thick] (10, 1) rectangle (14, 3) node[midway]{\LARGE Regulating  \\ \LARGE power bid};
  \node at (12,0) { \LARGE $\lambda_{h,\omega}^{\text{bid}}$};
  \node at (12,-2) {\LARGE D$-$1};
  \node at (12,-2.7) {\large (5:00pm)};

  \draw[rounded corners, fill=RectangleColor, very thick] (15, 1) rectangle (19, 3) node[midway]{\LARGE Real time};
  \node at (17,0) {\LARGE $\Gamma_{h,\omega}$};
  % {\LARGE $\Bigl( \bm{p}^{\rm{r},\uparrow}, \bm{\lambda}_{\omega}^{\text{bid}}, \bm{p}_{\omega}^{b,\uparrow}, \bm{p}_{\omega}^{b,\downarrow}, \bm{s}_{\omega}, \bm{T}_{\omega}^{c}, \bm{T}_{\omega}^{f}, \notag \\ \quad & \quad \bm{T}_{\omega}^{c, B}, \bm{T}_{\omega}^{f,B}, \bm{\phi}_{\omega}, \bm{g}_{\omega} \Bigr)$};
  \node at (17,-2) {\LARGE D};
  \node at (17,-2.7) {\large (12pm -- 12am)};

  % Right arrow
  \node (source) at (0, -1) {};
  \node (destination) at (24,-1){};
  \draw[->, line width=0.65mm, black](source)--(destination);

\end{tikzpicture}

\end{document}