### 2024-03-07 tex/highlights/highlights.tex

\documentclass{article}

\begin{document}

\section*{Highlights}

\begin{itemize}
    \item By creating a two-stage stochastic optimization model, TCLs can in tertiary reserves (mFRR) in the Nordics to exploit their flexibility potential.
    \item Load shifting provide an alternative revenue stream for TCLs which is not necessarily better for the power for the grid.
    \item Load shifting can be more appealing than mFRR for TCLs and has lower barriers to entry while being almost as profitable.
\end{itemize}

\end{document}