### 2024-03-07 tex/figures/timeline_mfrr_variables_tikz.tex

\documentclass[crop,tikz]{standalone}
\usepackage[usenames]{color} %used for font color
\usepackage{amssymb} %maths
\usepackage{amsmath} %maths
\usepackage[utf8]{inputenc} %useful to type directly diacritic characters
\usepackage{tikz}
\usetikzlibrary{calc,arrows,patterns,intersections}
\usetikzlibrary{shapes.multipart}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}
\begin{document}
\definecolor{RectangleColor}{RGB}{197,224,180}

\begin{tikzpicture}[scale=0.75,every text node part/.style={align=center}]

  \node[align=center, draw=none, fill=none, rotate=90] at (-0.5,4) {mFRR market};
  \node[align=center, draw=none, fill=none, rotate=90] at (-0.5,0) {Load shifting};

  \draw[rounded corners, fill=RectangleColor, very thick] (0, 3) rectangle (4, 5) node[midway]{\Large mFRR \\ \Large bidding};
  \node at (2,2) {\Large  $p_{h}^{\rm{r},\uparrow}$};
  \draw[dashed, black, line width=1.5pt] (4.5,-1) -- (4.5,5);
  \node at (4.5,-2) {\Large D$-$1};
  \node at (4.5,-2.7) {\large (9:30am)};
  \node[align=center, draw=none, fill=none, rotate=90] at (4.8,1.2) {mFRR gate closure};

  \draw[dashed, black, line width=1.5pt] (7,-1) -- (7,5);
  \node at (7,-2) {\Large D$-$1};
  \node at (7,-2.7) {\large (12:00pm)};
  \node[align=center, draw=none, fill=none, rotate=90] at (6.7,1.9) {Day-ahead market clearing};
  \draw[rounded corners, fill=RectangleColor, very thick] (7.375, 1) rectangle (11.375, 3) node[midway]{\Large Load shifting};
  \node at (9.375,0) {\Large  $p_{h}^{\text{Base}}$};

  \draw[rounded corners, fill=RectangleColor, very thick] (12, 3) rectangle (16, 5) node[midway]{\Large Regulating  \\ \Large power \\ \Large bidding};
  \node at (14,2) { \Large $\lambda_{h,\omega}^{\text{bid}}$};
  \draw[dashed, black, line width=1.5pt] (16.25,-1) -- (16.25,5);
  \node[align=center, draw=none, fill=none, rotate=90] at (16.65,1.9) {Regulation bid gate closure};
  \node at (16.25,-2) {\Large D$-$1};
  \node at (16.25,-2.7) {\large (5:00pm)};

  \draw[rounded corners, fill=RectangleColor, very thick] (17.5, 3) rectangle (21.5, 5) node[midway]{\Large Activation};
  \node at (19.5,2) {\Large $\Gamma_{h,\omega}$};
  % {\Large $\Bigl( \bm{p}^{\rm{r},\uparrow}, \bm{\lambda}_{\omega}^{\text{bid}}, \bm{p}_{\omega}^{b,\uparrow}, \bm{p}_{\omega}^{b,\downarrow}, \bm{s}_{\omega}, \bm{T}_{\omega}^{c}, \bm{T}_{\omega}^{f}, \notag \\ \quad & \quad \bm{T}_{\omega}^{c, B}, \bm{T}_{\omega}^{f,B}, \bm{\phi}_{\omega}, \bm{g}_{\omega} \Bigr)$};
  \node at (19.5,-2) {\Large D};
  \node at (19.5,-2.7) {\large (12pm -- 12am)};

  % Right arrow
  \node (source) at (0, -1) {};
  \node (destination) at (24,-1){};
  \draw[->, line width=0.65mm, black](source)--(destination);

\end{tikzpicture}

\end{document}